---
title: "Test your R install!"
subtitle: "Not graded"
output: html_document
---

# Introduction

This is an R install test document.  We compile it in a way that the
code will not be visible, you will just have the output.  Imagine this
is a report, and your boss is deeply uninterested in the code.  But
she very much wants to see the results!

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(echo = FALSE)
```

The document should compile without errors, and produce a nice html.


# Load data from a non-default package

```{r cars}
data(father.son, package="UsingR")
```

Here we loaded data (Galton 1886 parents/children height data) from
package _UsingR_.  You probably have to install that package.  The
dataset has `r nrow(father.son)` rows.

# Use **dplyr** commands

```{r}
fs <- father.son %>% 
  mutate(fheight=fheight*2.54,
         sheight=sheight*2.54)
```
Here we use dplyr to convert inches to centimeters: the fathers'
average height is `r round(mean(fs$fheight), 1)` cm.

If it works, you have a functioning dplyr installation.


# Do some plotting

Finally, here is the plot of the data (`message=FALSE` avoids
informative messages from code):

```{r, message=FALSE, fig.cap="Height of parents and offsprings"}
ggplot(fs, aes(fheight, sheight)) +
  geom_abline(slope=1, intercept=0, color="orangered3") +
  geom_point() +
  geom_smooth(method="lm", se=FALSE) +
  labs(x = "Fathers height (cm)",
       y = "Sons height (cm)")
```

The black dots depict parent-child pairs (although the data claims it
is about fathers and sons, Galton was in fact also using mothers and
daughters).  The red line is one-to-one relationship, meaning that
one cm taller parents have in average one cm taller children.  As the
blue regression line shows, this is not at all the case.  Taller
parents have taller children, but not as much above the mean as
parents.  This is what Galton (1886) called "regression to mediocrity"
and we call "regression to mean".
