---
title: "week3A"
author: "Lipsa Jena"
date: "10/12/2020"
output: html_document
---

#first read data
```{r}
gsa <- read.delim("../data/GSAF5.csv.bz2", stringsAsFactors = FALSE)
names(gsa)
```

#to insert a r code chunk in cmnd + option + i
# you can have seperate set directory in rmd file and in console

#below is traditional filtering
```{r}
gsa[gsa$Year == 2000,]
```

#pipes~ is %>% , it takes a result and feed it to another
```{r}
add(2*eggs, sugar) %>%
  whisk() %>%
```

# pipe operator 
# traditional approach
```{r}
print(sqrt(2))
```
```{r, message=False}
library(dplyr)
```


```{r, message=False}
2 %>%  sqrt() %>% print()

```
# exercise
compute and print mean of 1,2,3,4 using pipes
```{r}
c(1,2,3,4) %>% mean() %>% print() #first way
1:4 %>% mean() %>% print() #second way

```

# dplyr library

* select: select _variables_
* filter: filter _observations_
* mutate: compute on observation basis
* arrange: order
* summarize: collapse many obseravtions into a few figures

## select variables
```{r}

gsa %>% 
  select(Year, Country, Type, Fatal..Y.N.) %>%
  head(10)
```
```{r}
gsa %>%
  filter(Country == "AUSTRALIA") %>%
  select(Date:Country, Fatal..Y.N.) %>%
  head(10)
```
#filter country china and taiwan
```{r}
gsa %>%
  filter(Country %in%  c("CHINA","TAIWAN"))  %>%
  ## filter(Country ==  c("CHINA","TAIWAN"))  %>% checks only first few rows not the full dataset
  select(Date:Country, Fatal..Y.N.) %>%
  print()
```
## Arrange

```{r}
gsa %>%
  select(Date:Country, Fatal..Y.N.) %>%
  filter(!(Country %in% c("CHINA","TAIWAN"))) %>%
  filter(!is.na(Year)) %>%
  arrange(Year) %>%
  head(10)
```

## Arrange

```{r}
gsa %>%
  select(Date:Country, Fatal..Y.N.) %>%
  filter(!(Country %in% c("CHINA","TAIWAN"))) %>%
  filter(!is.na(Year)) %>%
  arrange(desc(Year), Country) %>%
  head(10)
```


## summarize
```{r}
gsa %>%
  filter(!is.na(Year)) %>%
  summarise(min = min(Year), avg = mean(Year), max = max(Year))
```

##grouped operations

* split data into groups
* imagine groups are independent
* summarize puts these summaries together
```{r}
gsa %>%
  filter(!is.na(Year)) %>%
  filter(Country != "") %>%
  group_by(Country) %>%
  summarise(min = min(Year)) %>%
  arrange(Country) %>% ## put string factor as false
  head(10)
```
