---
title: "inc07A"
author: "Lipsa Jena"
date: "11/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data(iris)
```

```{r}
head(iris)
dim(iris)
```

```{r}
library(tidyverse)
ggplot(iris, aes(Sepal.Length, Sepal.Width, col = Species))+
  geom_jitter(size = 3)
```

```{r}
setosa <- iris %>%
  filter(Species == "setosa")
dim(setosa)
```
# plot setosa only

```{r}
ggplot(setosa, aes(Sepal.Length, Sepal.Width))+
  geom_jitter(size = 2) +
  geom_smooth(method = "lm", se = FALSE) ## se is false. dont ponit outsiders
#lm is linear model, linear regression

```


Linear model y = B0 + B1.xi +ei | xi is explanatory variable, B0 and B1 aew numberds, y is dependent variable ; i is index of observation

ei is called epsilon i; our model has two components. 

**linear model is the deterministic components**
**ei is stochatic model; thats what moves the point away from the linear model**

beta null and beta one; estimated value.; yi = B0 +B1xi; **estimated values** and **residuals**
**predicted value**

* sum of squared errors (sse)
manipulate intercept and slope

-- linear model and linear reg -> diff; prof calls linear model is same as linear reg in terms of R

y = mx+c -> compare the linear part

## --------
from 8.1.2

```{r}
virginica <- iris %>%
  filter(Species == "virginica")
dim(setosa)
```

```{r}
beta0 <- 0
beta1 <-0.5
yHat <- beta0 +beta1*virginica$Petal.Length
SSE <- sum((virginica$Petal.Width-yHat)^2)
SSE
```


## estimate OLS model - ordinary least square

```{r}
m <- lm(Sepal.Width  ~  Sepal.Length, data = setosa)
summary(m)
```
B1 = 0.8

1f length is larger by 1 them width is by 0.8


## income and education

estimate 

```{r}
B1 = 5000
B0 = 10000 #(intercept)

```




