---
title: "Lab07"
author: "Lipsa Jena"
date: "11/18/2020"
output: html_document
---

```{r, include=FALSE}
library(ggplot2)
```

## 1. Diamonds price

### 1.1 Diamonds data

Getting the diamond dataset.

#### 1. Load the data and perform basic sanity checks. How many diamonds do we have?
```{r}

diamonds <- read.delim("../data/diamonds.csv.bz2")
```
**basic sanity check:**

checking the summary of diamonds data set:

```{r}
summary(diamonds)
```

Checking the first few rows of diamonds data set:

```{r}
head(diamonds)
```

**The total number of diamonds in the set are `r nrow(diamonds)`**

#### 2. Inspect the variable cut. What kind of different cuts are there? How frequent are those?

There are a total of 5 cuts and there frequencies can be seen below:

```{r}
table(diamonds$cut)
```

### 1.2 Regression analysis

#### 1. Include two variables: the mass of diamonds (carat) and cut. Estimate model of a form

```{r}
m <- lm(price ~ carat + cut, data = diamonds)
summary(m)
```

#### 2. What is the reference category for cut?

The reference categorty for 'cut' is Fair. As we can see in the above resultt set, that cutFair is only missing value in coeficients making it the reference and intercept.

#### 3. Interpret the following coeficients:

#### (a) What is "carat"?

It is `7871.08`. This means that for every 1 carat increase in the mass, the price of diamond increases by `7871.08`

#### (b) What is "cutIdeal"?

It is `1800.92`. This mean for every 1 increase in cutIdeal count, the price of the diamond increases by `1800.92`

#### (c) How much more expensive are ideal-cut diamonds compared to very good-cut diamonds (in average)?

The ideal cut diamonds are expensive by `r 1800.92 - 1510.14` when compared to very good-cut diamonds.

#### (d) What would be the (average) price for 1ct premium cut diamond?

The price of 1ct premium cut diamond is `r 7871.08 + 1439.08 -3875.47`


## 2. Challenge

### 1. Plot the price versus mass and add the regression line to it.


```{r}
ggplot(diamonds, aes(price, carat))+
  geom_jitter(size = 1) +
  geom_smooth()
```


### 2. Experiment with log scale for a) price, b) carat, c) both. Which plot does look the best?

a) log of price
```{r}
ggplot(diamonds, aes(log(price), carat))+
  geom_jitter(size = 1) +
  geom_smooth()
```


b) log of carat
```{r}
ggplot(diamonds, aes(price, log(carat)))+
  geom_jitter(size = 1) +
  geom_smooth()
```

c) log of both
```{r}
ggplot(diamonds, aes(log(price), log(carat)))+
  geom_jitter(size = 1) +
  geom_smooth()
```


According to me, log of both looks the best.

### 3. Add the corresponding log-transform to your model. You can run log-transformed models just like lm(log(price) log(carat))

```{r}
ggplot(diamonds, aes(log(price), log(carat)))+
  geom_jitter(size = 1) +
  geom_smooth()
```


```{r}
plot(lm(log(price) ~ log(carat), data = diamonds))
```





