---
title: "PS06"
author: "Lipsa Jena"
date: "11/19/2020"
output: html_document
---


```{r, include=FALSE}
library(ggplot2)
```


## 1 Housing Values in Boston

This dataset contains information about median house value for 506 neighborhoods in Boston, MA. It is collected in 1970s, in the good ol’ days when house prices were less than $50k.


### 1. Describe the data and variables that are part of the Boston dataset. Are there any missings? Any unreasonable values? Clean data as necessary

```{r}
boston_hd <- read.delim("../data/boston.csv.bz2")
```

Boston  neighborhood household data has a total of `r nrow(boston_hd)` rows. 

The variables are:

* **crim** : per capita crime rate by heighborhood.
* **zn** : proportion of residential land zoned for lots over 25,000 sq.ft.
* **indus** : proportion of non-retail business acres per neighborhood.
* **chas** : Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).
* **nox** : nitrogen oxides concentration (parts per 10 million).
* **rm** : average number of rooms per dwelling.
* **age** : proportion of owner-occupied units built prior to 1940.
* **dis** : weighted mean of distances to five Boston employment centres.
* **rad** : index of accessibility to radial highways.
* **tax** : full-value property-tax rate per $10,000.
* **ptratio** : pupil-teacher ratio by neighborhood.
* **black** : 1000(Bk − 0.63)^2 where Bk is the proportion of blacks by neighborhood.
* **lstat** : lower status of the population (percent).
* **medv** : median value of owner-occupied homes in $1000s.

Descriprion of data:
The total number of na values are `r sum(is.na(boston_hd))` rows.
Few of the unresonable values are:
1) For example the median value of owner is capped at 50k.
2) I dont see the importance of mentioning blacks by neighborhood.

Clean data:
The data looks clean to me, I dont see the need to clean data.


### 2. Use the following predictors: rm, lstat, indus, nox and dis

### 1) medv vs rm

#### 1.a) Make a scatterplot that displays how medv is related to that predictor and add regression line to that plot. Comment the result: do you see any relationship? Anything else interesting you see?

```{r}
ggplot(boston_hd, aes(rm, medv))+
  geom_jitter(size = 1, color = "darkred") +
  geom_smooth(method = "lm") 
```

I notice a positive relation in between rm and medv.

#### 1.b) Fit a simple linear regression model to predict the response. Interpret the slope (the effect of your explanatory variable). Is it statistically significant?

```{r}
m1 <- lm(medv ~ rm, data = boston_hd)
summary(m1)
```

The funtion for the medv value is f(rm)=-34.67+9.10rm. The slope and intercept in the function are statistically significant because both p-values are lesser than 0.05. We can reject the null hypothesis as there is no relationship between the two variables.


#### 1.c) Explain why do you think you see (or don’t see) the relationship on the figure/model.

The relationship between rm and medv is positive. 
In simple words, the median income tends to increase as the average number of rooms in a household increase. In my opinion, this could be due to the house with higher number of rooms are expensive when compared to houses with lower number of rooms.

### 2) medv vs lstat

#### 2.a) Make a scatterplot that displays how medv is related to that predictor and add regression line to that plot. Comment the result: do you see any relationship? Anything else interesting you see?

```{r}
ggplot(boston_hd, aes(lstat, medv))+
  geom_jitter(size = 1 , color = "darkgreen") +
  geom_smooth(method = "lm")
```

I see a negetive relation in lstat (lower status of the population) and medv.


#### 2.b) Fit a simple linear regression model to predict the response. Interpret the slope (the effect of your explanatory variable). Is it statistically significant?

```{r}
m1 <- lm(medv ~ lstat, data = boston_hd)
summary(m1)
```

The funtion for the medv value is f(lstat)=+34.55-0.95lstat. The slope and intercept in the function are statistically significant because both p-values are lesser than 0.05. We can reject the null hypothesis as there is no relationship between the two variables.

#### 2.c) Explain why do you think you see (or don’t see) the relationship on the figure/model.

The relationship between lstat and medv is negetive. 
In simple words, the median income tends to decrease as the lower status of populations increase. In my opinion, this could be due to the housings which are expensive are in neighborhoods where the lower status of people is low. In otherwoors, areas with lower status have cheaper housing opyions whereas the areas with higher status have expensive housings options.

### 3) medv vs indus

#### 3.a) Make a scatterplot that displays how medv is related to that predictor and add regression line to that plot. Comment the result: do you see any relationship? Anything else interesting you see?

```{r}
ggplot(boston_hd, aes(indus, medv))+
  geom_jitter(size = 1, color = "darkred") +
  geom_smooth(method = "lm") 
```

I see a negetive relationship between indus and medv

#### 3.b) Fit a simple linear regression model to predict the response. Interpret the slope (the effect of your explanatory variable). Is it statistically significant?

```{r}
m1 <- lm(medv ~ indus, data = boston_hd)
summary(m1)
```

The funtion for the medv value is f(indus)=+29.75-0.64indus. The slope and intercept in the function are statistically significant because both p-values are lesser than 0.05. We can reject the null hypothesis as there is no relationship between the two variables.

#### 3.c) Explain why do you think you see (or don’t see) the relationship on the figure/model.

The relationship between indus and medv is negetive. 
In simple words, the median income tends to decrease as the proportion of non-retail business acress increase. In my opinion, this could be due to the housings which are expensive are away from industrial areas and cheaper houses are available closer to industrial area. 

### 4) medv vs nox

#### 4.a) Make a scatterplot that displays how medv is related to that predictor and add regression line to that plot. Comment the result: do you see any relationship? Anything else interesting you see?

```{r}
ggplot(boston_hd, aes(nox, medv))+
  geom_jitter(size = 1, color = "darkgreen") +
  geom_smooth(method = "lm")
```

I see a negetive relation between nox and medv

#### 4.b) Fit a simple linear regression model to predict the response. Interpret the slope (the effect of your explanatory variable). Is it statistically significant?

```{r}
m1 <- lm(medv ~ nox, data = boston_hd)
summary(m1)
```

The funtion for the medv value is f(nox)=+41.34-33.91nox. The slope and intercept in the function are statistically significant because both p-values are lesser than 0.05. We can reject the null hypothesis as there is no relationship between the two variables.

#### 4.c) Explain why do you think you see (or don’t see) the relationship on the figure/model.

The relationship between indus and nox is negetive. 
In simple words, the median income tends to decrease as the nitrogen oxides concentration increase. In my opinion, this could be due to the housings which are expensive have lesser polution as they are away from boston work as well as industrial areas. On the other hand, cheaper housing options are available in the localities which are closer to Boston city area, making it more pollution prone.

### 5) medv vs dis

#### 5.a) Make a scatterplot that displays how medv is related to that predictor and add regression line to that plot. Comment the result: do you see any relationship? Anything else interesting you see?

```{r}
ggplot(boston_hd, aes(dis, medv))+
  geom_jitter(size = 1, color = "darkred") +
  geom_smooth(method = "lm")
```

I see a positive relation between medv and dis.

#### 5.b) Fit a simple linear regression model to predict the response. Interpret the slope (the effect of your explanatory variable). Is it statistically significant?

```{r}
m1 <- lm(medv ~ dis, data = boston_hd)
summary(m1)
```

The funtion for the medv value is f(dis)=+18.39+1.09dis. The slpoe and intercept in the function is statistically significant because both p-values are lesser than 0.05. We can reject the null hypothesis as there is no relationship between the two variables.

#### 5.c) Explain why do you think you see (or don’t see) the relationship on the figure/model.

The relationship between dis and medv is positive. 
In simple words, the median income tends to increase as the average distances to Boston employment centers increase. In my opinion, this could be due to the number of rooms being higher on the outskirts and areas closer to Boston might have 1 bedroom houses. As we have seen in this dataset, the houses with more rooms cost more.

### 3. Comment the results: are plots where you clearly can see a relationship related to models where the effect is statistically significant?

Yes, all of them are stastically significant as the p values are lesser than 0.05

In my opinion, rm, lstat, dis and nox have a clear scatter plot ouput. On the other hand, Indus has a very unclear scatter plot. Its difficult to set a relationship in case of Indus.

### 4. Fit a multiple regression model to predict the response using all the predictors. Describe your results. For which predictors can we reject the null hypothesis H0 : βj = 0?

```{r}
m1 <- lm(medv ~ ., data = boston_hd)
summary(m1)
```

We can not reject null hypothesis where p-value is greater that 0.05; In this scenario only **indus** and **age** are such scenario where p-value is greater than zero. Hence, we can not reject the Null hypothesis. All the other variables we can reject Null hypothesis, below is a list:

* crim
* zn
* chas
* nox
* rm
* dis
* rad
* tax
* ptratio
* black
* lstat


#### 5. Interpret the results for rm, lstat and indus. Are the results statistically significant?

Interpret:
rm - With every 1 unit increase in rm (with everything else constant), the medv value increases by 3.810e+00.
lstat - With every 1 unit increase in lstat (with everything else constant), the medv value decrease by 5.248e-01.
indus - With every 1 unit increase in indus (with everything else constant), the medv value increases by 2.056e-02.

Looking at the R output we can say that rm and lstat are statistically significant whereas indus is statistically insignificant.

#### 6. How do your results from 2 compare to your results from 4? Compare the results for those predictors you used for simple regression above. Explain why do the values differ. Do they still tell the same basic story?

When compared to 2 and 4: 
* indus was statistically significant in simple regression but in multiple regression it becomes insignificant. 
* rm and lstat were statistically significant in both simple regression and multiple regression. 

In simple regression, we compared them independently with median values. In multiple regression 
median values are influnced by neibourhood and other factors. 

In my opinion, the output are still similar excpet two factors - indus and age, it still provides the same pictire.

## 2 Interpret Regression Results

### 1. Do neighborhoods with more evictions see more or less 311 calls? By how much?

Yes, the neighborhoods with more evictions see more 311 calls. The number of 311 calls increases by 0.048 with a unit increase in No. of evictions

### 2. Is the figure statistically significant (at 5% level)?

Yes, for the number of evictions, the p value is lesser than 0.01 which makes it statistically significant.

### 3. How is poverty rate associated with 311 calls? How much more (or less) calls there are in neighborhoods with 10 pct point more poverty?

Poverty value is negetively associated with 311 calls. 311 calls would decrease by 0.14 with a 1 unit increase in poverty in the total population. 
**so, for a 10 percent increase in poverty the 311 calls would decrease by 1.4**

### 4. What can you tell about association of race (white) and calls?

White(race) is negetively associated with 311 calls. 311 calls would decrease by 0.038 with a 1 unit increase in White proportion in the total population

### 5. Is older median age associated with more or less 311 calls? At which level is this statistically significant?

Yes, older median age is associated with more 311 calls
Median age is positively related to 311 calls, which means the number of 311 calls would increase by 0.0067 with a unit increase in median age.

At 1% level it becomes statistically significant.

Here the p value is lesser than 0.01, which makes it statistically significant.

### 6. The value for housing density is -0.13. What does this number mean?

The housing density is negetively associated with 311 calls. 311 calls would decrease by 0.13 with a 1 percent increase in housing density. 

### 7. The omitted category for city is Austin, TX. Are there more or less calls in similar neighborhoods in Philadelphia, compared to Austin? By how much?

Neighborhoods in Philadelphia has lesser 311 calls when compared with Austin texas.
Philadelphia 311 calls decreases by 0.56 when compared to Austin


## Extra Credit


### 1. Do neighborhoods with more evictions see more or less 311 calls? By how much?

__It's a log - log relation ship here.__

Yes, the neighborhoods with more evictions see more 311 calls. The number of 311 calls increases by 0.0086 percent with a 1 percent increase in No. of evictions

### 2. Is the figure statistically significant (at 5% level)?

Yes, for the number of evictions, the p value is lesser than 0.01 which makes it statistically significant.

### 3. How is poverty rate associated with 311 calls? How much more (or less) calls there are in neighborhoods with 10 pct point more poverty?

__It's a log - linear relation ship here.__

Poverty value is negetively associated with 311 calls. 311 calls would decrease by 0.057 with a 1 percent increase in poverty in the total population. 

**so, for a 10 percent increase in poverty the 311 calls would increase by 0.57**

### 4. What can you tell about association of race (white) and calls?

__It's a log - linear relation ship here.__

White(race) is negatively associated with 311 calls. 311 calls would decrease by 0.026 with a 1 percent increase in White proportion in the total population

### 5. Is older median age associated with more or less 311 calls? At which level is this statistically significant?

__It's a log - linear relation ship here.__

Yes, older median age is associated with more 311 calls
Median age is positively related to 311 calls, which means the number of 311 calls would increase by 0.0010 with one percent increase in median age.

At 1% level it becomes statistically significant.

Here the p value is lesser than 0.01, which makes it statistically significant.

### 6. The value for housing density is -0.13. What does this number mean?

__It's a log - log relation ship here.__

The housing density is negatively associated with 311 calls. 311 calls would decrease by 0.012 percent with a 1 percent increase in housing density. 

### 7. The omitted category for city is Austin, TX. Are there more or less calls in similar neighborhoods in Philadelphia, compared to Austin? By how much?

__It's a log - linear relation ship here.__

Neighborhoods in Philadelphia has lesser 311 calls when compared with Austin texas.
Philadelphia 311 calls decrease by 0.04 percent when compared to Austin


## Time spent

It took me a total of 25+ hours. I spent a lot of time reading the materials provided.