---
title: "PS05"
author: "Lipsa Jena"
date: "11/12/2020"
output: html_document
---
## 1 Are Sons Taller than Fathers?
### 1.1 Descriptive Analysis
#### 1. load the fatherson.csv data. Do the basic descriptive work on it:
#### what is the number of observations? Are there any missings?


```{r,include=FALSE}
# Loaded the necessary data

library(tidyverse)
```


```{r}
fsdata <- read.delim("../data/fatherson.csv.bz2")
total <- nrow(fsdata)

na_rowsf <- sum(is.na(fsdata$fheight))
na_rowss <- sum(is.na(fsdata$sheight))

```

The “fatherson.csv” is now stored in variable `fsdata`.
Some basic inspection of data:

a) total number of observations are `r total`.
b) total number of rows with fathers height missing values are `r na_rowsf`.
b) total number of rows with fathers height missing values are `r na_rowss`.

#### 2. Describe fathers and sons: compute the mean, median, standard deviation, and range of their heights. 


```{r}
meanF <- mean(fsdata$fheight)
medianF <- median(fsdata$fheight)
sdF <- sd(fsdata$fheight)
rangeF <- range(fsdata$fheight)
```

a) The values for fathers height are:
* Mean: `r meanF`
* Median: `r medianF`
* Standard deviation: `r sdF`
* Range of the heights: `r rangeF`


```{r}
meanS <- mean(fsdata$sheight)
medianS <- median(fsdata$sheight)
sdS <- sd(fsdata$sheight)
rangeS <- range(fsdata$sheight)
```

b) The values for sons height are:
* Mean: `r meanS`
* Median: `r medianS`
* Standard deviation: `r sdS`
* Range of the heights: `r rangeS`

#### According to these figures, who are taller: fathers or sons?

The mean of fathers heights are `r meanF` and the mean of sons height are `r meanS`. hence, sons mean height is `r meanS - meanF` cm greater than fathers mean height.
The median of fathers heights are `r medianF` and the median of sons height are `r medianS`. hence, sons median height is `r medianS - medianF` cm greater than fathers median height.

**Sons are taller than their father**

#### 3. Lets add a graphical comparison. Create density plots of both heights on the same figure. Comment the density plots. Which distribution do these resemble? Do they agree with the conclusion above that sons are taller?


```{r}
fsdata_plot <- fsdata %>% 
  pivot_longer(c("fheight","sheight"), names_to = "fheight", values_to = "sheight") %>%
  select(category=fheight, height=sheight)

```

Density plot:

```{r}

ggplot(fsdata_plot, aes(x=height, color = category)) + 
  geom_density(alpha=0.4) +
  scale_color_manual(name="category", 
                        labels = c("Fathers height", 
                                   "Sons height"), 
                        values = c("fheight"="blue", 
                                   "sheight"="green"))

```

#### 4. Finally, for the further reference, compute how much taller are sons in average.

The mean of fathers heights are `r meanF` and the mean of sons height are `r meanS`. hence, sons mean height is `r meanS - meanF` cm greater than fathers mean height.

### 1.2 Monte Carlo approach

#### 1. compute the overall mean and standard deviation of pooled fathers’ and sons’ heights. (I.e combine all heights, and compute just one mean and one standard deviation for this combined data.)

```{r}
meanA <- mean(fsdata_plot$height)
sdA <- sd(fsdata_plot$height)

```

The values for both the height are:

* Overall Mean: `r meanA`
* Overall Standard deviation: `r sdA`

#### 2. now create two sets of random normals, both with the same mean and standard deviation that you just computed above. Call one of these “fake fathers” and the others “fake sons”.

```{r}
fakefathers <- rnorm(total, mean=meanA, sd=sdA)
fakesons <- rnorm(total, mean=meanA, sd=sdA)
```

#### What is the fake father-fake son mean difference? Compare the result with that you found in the previous problem

```{r}
diff_F <- mean(fakesons) - mean(fakefathers)
```

The fake father and fake son mean difference is `r diff_F`cm. 
This is smaller than the actual difference of 2.5 cm

#### 3. Now repeat the previous question a large number of times R (1000 or more). Each time store the difference, so you end up with R different values for the difference.

```{r}
sim <- function(x) {   
  fakefathers <- rnorm(total, mean=meanA, sd=sdA)
  fakesons <- rnorm(total, mean=meanA, sd=sdA)  
  mean(fakesons) - mean(fakefathers) 
}

diffs <- sapply(1:1000, sim)
```

I repeated the function 1000 times and all the differences are stored in variable `diffs`.

#### 4. What is the mean of the differences? Explain why do you get what you get.

```{r}
cat("The mean of the differences is: ", mean(diffs))
```

As expected the mean value is closer to zero. This becomes closer to zero due to central limit theoram (CLT). The distribution of sample approximates a normal distribution - “bell curve” as the sample size becomes larger.

#### 5. What is the standard deviation of the differences?

```{r}
cat("The standard deviation of the differences is: ", sd(diffs)) 
```

#### 6. What is the largest value among the differences (in absolute value)? How does it compare to the actual sons/fathers difference of 2.5.

```{r}
cat("The largest value among the differences is: ", max(abs(diffs)))  
```

The largest value among the differences : `r max(abs(diffs))` is smaller than the actual fathers and sons difference which is 2.5 cm 

#### 7. find 95% confidence intercals for the differences you computed.Does the actual difference fall inside or outside of the CI?

```{r}
cat("The 95% confidence intercals for the differences computed is: \n")
quantile(diffs,c(0.025, 0.975))
```

The actual difference is outside of the CI.

#### 8. What is your conclusion? Can you confidently say that sons are taller than fathers? Why?

The actual sons/fathers difference is located outside of the 95% confidence intervals. Thus, we should reject H0 and conclude that sons are taller than fathers.

### 1.3 t-test 
Simulations are great but also cumbersome. t-test comes to help.

#### 1. Compute the standard error for the difference of means.

```{r}
SE <- sqrt(sdF^2/total + sdS^2/total)

```
The standard error for the difference of means is `r SE`.

#### 2. Compute the t-value (OIS denotes it by T). Here we ask to compute it yourself, not use any pre-existing t-test functions! What is the t-value you find?

```{r}
point_estimate = meanS - meanF
t_value <- point_estimate/SE

```
The calculated t-value is `r t_value`.

#### 3. Look up the t-distribution table. What is the likelihood that such a t-value happens just by random chance?

```{r}
df <- total -1
t.test(fsdata$sheight, fsdata$fheight)
```

our calculated degree of freedom in this scenario is `r df` > 30 . The larger the degrees of freedom, the more closely the t-distribution resembles the standard normal distribution.

The likelihood is less than 0.05 % as the t-value is greater than 2.58, the t-value of which is created by top 0.5% of the distribution and `r df`.


#### 4. finally, state clearly your conclusion. Is the actual difference you see compatible with H0 that fathers and sons are of similar height?

The t-value is located far right tail in the distribution, not at all compatible with H0. We should reject H0 and conclude that fathers and sons are not of similar height.


## 2. Extra credit: parallel computing

#### 1. Lets return to your MC example. Increase the number of repetitions R substantially, so it runs for at least 10 seconds.

```{r}
diffs_ec <- sapply(1:100000, sim)
```


#### 2. Conduct the MC analysis using a parallel loop.

```{r ,include=FALSE}
library(foreach)
library(doParallel)
library(knitr)
```


```{r}
#setup parallel backend to use many processors
cores=detectCores()
cl <- makeCluster(2, setup_strategy = "sequential") #not to overload your computer
registerDoParallel(cl)

diffs_pp <- foreach(i=1:100000) %dopar% {
   diff_pp = sim() #calling a function
}
#stop cluster
stopCluster(cl)
```

references:

* https://riptutorial.com/r/example/5164/parallel-processing-with-foreach-package
* https://github.com/rstudio/rstudio/issues/6692


#### 3. Time your code. Create a table that shows how the simulation time depends on the number of employed CPU cores. Can you get a noticeable speed improvement by running the simulation code in parallel?

```{r}
cores <- c(1,2,3,4)
time_duration <- c("59 s","35.9 s","30.3 s","29.2 s")

tab <- data.frame(cores, time_duration)
```

I ran the the code under question two with increasing core one by one. With every increase the processing time decreased.

```{r}
kable(tab)
```

#### 4 Finally, use the fastest approach and increase R as much as your computer can stand. Sure, you can afford to run it for 10 seconds. Maybe you can afford an hour. Report how large R did you use and how long time did it take. See how large difference you can find. Can you get anything comparable to the actual difference in the data?

```{r}
#diffs_lc <- sapply(1:1000000, sim)
#max(abs(diffs_lc))
#range(diffs_lc)
```

I have commented the above code as it takes really long time to knit with such a high value of R (set to 1000000).

I ran for 1 million times and it took 10 minutes 30 seconds to run. The largest values in this set was 1.260594.The range of 1 million differences lies between - 1.260594  and 1.250474.
