---
title: "IMT573 Lab 5: probability"
author: "Lipsa Jena"
date: "10/28/2020"
output: html_document
---

## Toss two coins

#### 1. Which objects are you tossing? If these are not coins, what event will you call "heads"?

I am tossing two coins.One fifty cents coin and one Quarter dollar coin.

#### 2. Toss two coins (or other objects) at least 20 times each. Each time note how many heads did you get.

I selected two coins and I am calling them `Gold Coin` and `Silver Coin`. I am assigning head as 1 and tail as 0, below is data for how it turned out:

```{r}
Gold_Coin <- c(0,0,1,1,0,0,1,1,0,1,1,0,1,0,1,1,0,0,0,0)
Silver_Coin <- c(1,1,1,1,0,1,0,1,1,0,0,0,0,1,1,1,1,0,0,0)
```

#### 3. Create a (readable)! table about the number of heads in each trial.

Creating a data frame - `coins.data` to store my result set of tossing coins

```{r}
Trail <- 1:20

#counting the total number of heads.
Number_of_heads <- Gold_Coin + Silver_Coin 

coins.data <- data.frame(Trail, Gold_Coin, Silver_Coin, Number_of_heads)

```

```{r,include=FALSE}
# Loaded the necessary data
library(dplyr)
```

Here presenting them in table format: 

```{r}
knitr::kable(coins.data)
```

#### 4 Show a summary table of your results (0 heads~ how many times; 1 head~how many times..)

```{r}
#count the number of times 0 heads appear:
head0 <- coins.data %>%
  filter(Number_of_heads == 0) %>%
  nrow()

#count the number of times 1 head appears:
head1 <- coins.data %>%
  filter(Number_of_heads == 1) %>%
  nrow()

#count the number of times 2 heads appear:
head2 <- coins.data %>%
  filter(Number_of_heads == 2) %>%
  nrow()

# Preparing to show summary data in table format.

cases <- c("The number of times 0 heads appear", "The number of times 1 head appears", "The number of times 2 heads appear")
Count_of_heads <- c(head0,head1,head2)
```

Summary table is :

```{r}
#creating a data frame to store summary table
summary_table <- data.frame(cases, Count_of_heads)

knitr::kable(summary_table)

```

#### 5. Compute mean of your results

Calculating the mean of total number of heads:

```{r}
meanC <- mean(coins.data$Number_of_heads)
cat("Mean of total number of heads is :", meanC)
```

#### 6. Compute variance of your results. We recommend to use the shortcut formula.

```{r}
# We have mean from previous question. Lets calculate mean^2

meanCsq <- meanC^2
cat("mean square is: ", meanCsq)

# Lets calculate the Square mean of all the heads.

sqMean <- sum(Number_of_heads^2)/20
cat("square mean is: ", sqMean)

variance <- sqMean-meanCsq

cat("Variance of Number of heads is: ", variance)


```

#### 7. Plot a histogram of your results.

Below is a plot of histogram:

```{r}
hist(coins.data$Number_of_heads, breaks = 10,  col = "pink")
```

#### 8. What type of distribution is this, discrete or continuous? Do you know how it is called?

The histogram above is a discrete distribution. Discrete distribution is one in which the data can only take certain fixed values. For example; here the variable `Number of heads` can have only three values; i.e, 0, 1, 2.
As the `Number of heads` cannot have values like 0.5 or 1.5, this is a **Discrete distribution**.

#### 9. Does it look symmetric, or right/left skewed?

This distrubution is perfectly symmetric. As professor mentioned in the class this is expected to be a normal distribution which is __Symmetric__.