---
title: "IMT573 Lab4: Get game scores from the web"
author: "Lipsa Jena"
date: "10/21/2020"
output: html_document
---

```{r, include=FALSE}
library(rvest)
library(dplyr)
```



## 1. Scrape recent college football scores


In this lab 4, we will bw working on scraping data from the web.


#### 1. Load and parse the college football page https://www.sports-reference.com/cfb/


Loading the website name and storing the html in the `coll_football` variable.

```{r}
url <- "https://www.sports-reference.com/cfb/"
coll_football <- read_html(url)
coll_football
```


#### 2. Find the table "Recent Games" a little down on the page. What is the element that contains the "Recent Games" text?

Recent game is in the **h2** element.


#### 3. This element is contained inside a DIV. What is the class of that DIV?

This element is inside **game_summaries**, class of the DIV.


#### 4. Extract that DIV. Call this "summaries".

```{r}
summaries <- coll_football %>%
  html_nodes("div.game_summaries") 

summaries
```


### 5. Underneath the title "Recent Games" you see a lot of scores. 

Yes.


#### 6. Inside of which elements are those texts embedded?

It is inside "game_summary.nohover".

```{r}

hover <- summaries %>% 
  html_nodes("div.game_summary.nohover") 

hover
  
```


```{r}

hover %>%
  html_nodes("table.teams") 

```


#### 7. Extract all those elements, and texts from therein. Print those texts.

```{r}
for(i in 1:length(hover)){
  cat("Game ", i, "\n")
  hover[[i]] %>% html_text() %>% gsub("[[:space:]]+", " ",.) %>% cat()
  cat("\n-----------\n")
}
```



## 2. Challenge


#### 1. create the data frame.

```{r}


df <- data.frame(Team_1=character(), 
                 Score_1 =character(),
                 Team_2=character(), 
                 Score_2 =character(),
                 stringsAsFactors=FALSE)
```


#### 2. create a list to store the extracted values.

```{r}
k <- list()

## for loop to store all the extracted value
for(i in 1:length(hover)){
  k[[i]] <- hover[[i]] %>% html_text() %>% gsub("[\t\n]+", "<>",.) %>% strsplit("<>")
  
  
}
```


#### 3. Moving the list to dataframe.

```{r}


for(i in 1:length(k)){
  df <- df %>% add_row(Team_1 = k[[i]][[1]][2],Score_1 = k[[i]][[1]][3],Team_2 = k[[i]][[1]][5],Score_2 = k[[i]][[1]][6])
  
}


```


#### 4. values in data frame:

```{r}
df
```


#### 5. Note for professor Ott and Bingbing:

Challenge was a good exercise, but I would like to know if you have a better solution. As my solution is quite cluttered. Also, I have a question regarding the cleaning the team names. for example how do I remove brackets and number from Georgia (3). As the bracket and number appears with only few name, so I couldnt use substr in the for loop.
