---
title: "INFX 573 Lab: Exploring Data"
author: "Lipsa Jena"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
urlcolor: blue
output: html_document
---

Collaborators: 

1. Aditya Challa

# Explore Titanic Data: Survival and Passenger Class

## Preparation

Load any add-on packages if needed:

```{r Setup, message=FALSE}
library(tidyverse)  # loaded tidyverse
```

## Explore Data

### 1. Sinking of RMS  Titanic analysis:
As part of this Assignment, we are going to analyse an important question of the RMS  Titanic sinking:
**Did higher class passangers have more chance to survive?** 

### 2. Load data

```{r Read in Data, eval=FALSE}
titanic <- read.csv("../data/titanic.csv.bz2")  ##loading data
                           
```

_we loaded the dataset into our R project._ 

### 3. Inspect data.

#### a) rows and columns

```{r inspect}
## inspecting data
titanic <- read.csv("../data/titanic.csv.bz2")
nrow(titanic)
ncol(titanic)
```

_The number of variables in this  are `r ncol(titanic)`  and the  number of cases are `r  nrow(titanic)`._

#### b) first and last lines

```{r}
head(titanic,2) 
tail(titanic,2)
```

_The **head** command shows top rows and the **tail** command shows from the bottom rows._

#### c) important variables
The  important variables for this analysis are: 'pclass'  and 'survived' which will  tell us the survival rate in titanic with respect to each class.

#### d) checking missing values in the relevent variables
```{r}
## checking if any of the relevent variables are missing values
sum(is.na(titanic$pclass)) 
sum(is.na(titanic$survived)) 
sum(is.na(titanic$sex)) 
sum(is.na(titanic$age)) 
```

### 4. How is survival probability related to passenger class?

```{r}
## split the data into classes
first <- titanic[titanic$pclass == 1,]
second <- titanic[titanic$pclass == 2,]
third <- titanic[titanic$pclass == 3,]

mean(first$survived, na.rm = TRUE) ## survival rate in first class
mean(second$survived, na.rm = TRUE) ## survival rate in second class
mean(third$survived, na.rm = TRUE) ## survival rate in third class

```

_Yes, as per our analysis its true that the higher class had a higher survival rate._


### 5. Why do you think this was the case?

_If we look at our data only from a class difference perspective, it looks like that the passanger from higher class had a better chance of survival, may be they were provided better facilities and safety. \n Lets look at this from a different perspective, here is survival of each sex type:_

```{r}

male <- titanic[titanic$sex == 'male',]
female <- titanic[titanic$sex == 'female',]

mean(male$survived, na.rm = TRUE) ## survival of male passengers
mean(female$survived, na.rm = TRUE) ## survival of female passengers
```

_The male survival rate is extremely lower than female passangers. That tells us that the female passangers were given first priority to escape irrespective of their seat class. I think the biggest learning of this assignment is to keep an open mind and inspect the other aspects of the provided dataset._



